---
title: "misc_equations"
author: "Elizabeth Zhang"
date: "`r Sys.Date()`"
geometry: margin=1.5in
output: pdf_document
header-includes:
  - \usepackage{amsmath}
---

```{r}
library(tidyverse)
library(kableExtra)
library(knitr)
```
\newpage

# kmr

Kernel machine regression defines the regression relationship using a flexible function $h: \mathbb{R}^M \rightarrow \mathbb{R}$, where

$$
Y_i = h(\textbf{x}_i) + \textbf{z}_i^\top \boldsymbol{\beta}_{\textbf{z}} + \varepsilon_i,
$$

\noindent Here, $Y_i$ is the outcome at a given point, $\textbf{x}_i=[x_1,\dots,x_M]^\top$ is a vector of $M$ chemicals,  $\textbf{z}_i$ and $\boldsymbol{\beta}_{\textbf{z}}$ are vectors of covariates and their weights, respectively, and $\varepsilon_i \overset{\mathrm{iid}}{\sim} N(0, \sigma^2)$. 

$h(\cdot)$ is obtained using the Gaussian kernel $k: \mathbb{R}^M\times \mathbb{R}^M \rightarrow \mathbb{R}$, defined as 

$$
k(\textbf{x}, \textbf{x}') = \textrm{exp}\bigg\{
-\frac{\sum_{m=1}^M(x_m-x_m')^2}{\rho} \bigg\},
$$

\noindent where $x$ are the predictor values of a second subject, and $\rho$ is a tuning parameter. 

\newpage

# bkmr priors

We define a weight, $r_m$, on each exposure by augmenting the kernel function as

$$
k(\textbf{x}, \textbf{x}'|\textbf{r}) = \text{exp}\bigg\{ -\sum_{m=1}^Mr_m(x_m-x_m')^2 \bigg\},
$$

where $r_m=1/{\rho_m}$ is the inverse of the tuning parameter $\rho_m$ for each $\textbf{x}_m$.

To allow $r_m$ to equal 0 with non-zero probability, we first define an indicator variable determining whether or not a predictor is included in the model, which is denoted and distributed as

$$
\delta_m \sim \text{Bernoulli}(\pi),
$$

\noindent where $\pi$ is the prior probability of inclusion. Now, we can assume a "slab-and-spike" prior on $r_m$, distributed as

$$
r_m|\delta_m \sim \delta_mf(r_m) + (1-\delta_m)P_0, 
$$

\noindent where $f(\cdot)$ is some pdf with support $\mathbb{R}^+$, and $P_0$ denotes the density with point mass at 0.

We define the following prior on $r_m$:

\begin{gather*}
\delta_m \sim \text{Bernoulli}(\pi),\text{ and}\\
r_m|\delta_m \sim \delta_mf(r_m) + (1-\delta_m)P_0, 
\end{gather*}

where $\pi$ is the prior probability of inclusion, $f(\cdot)$ is some pdf with support $\mathbb{R}^+$, and $P_0$ denotes the density with point mass at 0.

The posterior means of $\delta_m$ represent posterior inclusion probabilities (PIPs) of $\textbf{x}_m$, which can be used as measures of the relative importance of each exposure.

Posterior means of $\delta_m$ $\Rightarrow$ posterior inclusion probabilities (PIPs) of $\textbf{x}_m$.

\newpage

# spline

BSR uses spline regression to define the regression relationship as

$$
Y_i = f(\textbf{x}_i) + \textbf{z}_i^\top \boldsymbol{\beta}_{\textbf{z}} +\varepsilon_i,
$$

\noindent where $f$ is defined by a set of basis functions on the exposures, $\textbf{x}_i$, $\textbf{z}_i$ and $\boldsymbol{\beta}_{\textbf{z}}$ are the covariates and their associated weights, and $\varepsilon_i$ is a random variable where $\boldsymbol\varepsilon \overset{\mathrm{iid}}{\sim} N(0, \sigma^2)$.

BSR uses a natural spline regression. A general definition of the $K$ basis functions for a natural spline with interior knots $\xi_j$, $j=1,\dots,K$ over $K+1$ disjoint intervals is given by:

\begin{gather*}
b_1(X)=1, \hspace{0.25in} b_2(X)=X, \hspace{0.25in} b_{k+2}(X)=d_k(X)-d_{K-1}(X), \\
d_k(X)=\frac{(X-\xi_k)^3_+ - (X-\xi_K)^3_+}{\xi_K-\xi_k}.
\end{gather*}

\noindent Here, the regression model is defined as $f(X) = \sum_{j=1}^K\beta_jb_j(X)$.

\newpage

# bsr priors

BSR assumes the following general model formulation: 

\begin{gather*}
f(\textbf{x}_i)= \sum_{h=1}^Hf^{(h)}(\textbf{x}_i), \\
f^{(h)}(\textbf{x}_i)= \sum_{m_1=1}^M\widetilde{x}_{im_1}\boldsymbol\beta_{m_1}^{(h)} + 
\sum_{m_1=2}^M\sum_{m_2<m_1}\widetilde{x}_{im_1m_2}\boldsymbol\beta_{m_1m_2}^{(h)} + \dots,
\end{gather*}

\noindent where $\widetilde{X}_m=[b_{m1}(X_m),\dots,b_{md}(X_m)]$ represents a $d$-dimensional basis function expansion for the $m$th term, and $f^{(h)}(\textbf{x}_i)$ includes a summation of all $M$-way interactions. 

We define the following prior on $\boldsymbol\beta_S^{(h)}$:

\begin{align*}
P(\zeta_{mh}=1) &= \tau_h^{\zeta_{mh}}(1-\tau_h)^{1-\zeta_{mh}} 
I(A_h\not\subset A_{h'}\forall h'\neq h \textrm{ or } A_h=\{\}),\\
&\textrm{where } A_h=\{m:\zeta_h=1\},\textrm{ and} 
\end{align*}

\begin{align*}
P(\boldsymbol\beta_S^{(h)}|\boldsymbol\zeta) &= \bigg(1-\prod_{m\in S}\zeta_{mh}\bigg)P_{\textbf{0}} +
\bigg(\prod_{m\in S}\zeta_{mh}\bigg) \psi_1(\boldsymbol\beta_S^{(h)}), \\
&\textrm{where } S \textrm{ is some subset of } {1, 2, \dots, m},
\end{align*}

\noindent where $\zeta_{mh}$ have prior probability of inclusion $\tau_h$, $I()$ is an indicator to help with identifiability issues, $P_{\textbf{0}}$ denotes the density with point mass at $\textbf0$, and $\psi_1()$ is a multivariate normal distribution with mean $\textbf0$ and covariance $\boldsymbol\Sigma_{\boldsymbol\beta}$, a diagonal matrix with $\sigma^2\sigma_{\boldsymbol\beta}^2$ on the diagonals. 

# equations for interactions

$$
Y = x_1 + x_2
$$

$$
Y = x_1 + x_2 + 0.5(x_1*x_2)
$$

$$
Y = x_1 + x_2 + 0.2(x_1*(x_2-1)^2)
$$

$$
\widehat Y = \widehat{\beta}_0 + \widehat{\beta}_1x_1 + \widehat{\beta}_2x_2 + \widehat{\beta}_{12}x_1x_2
$$

# specification of interaction terms

```{r}
equations <- data.frame(
  type = c("Multiplicative", "Polynomial", 
           "Smaller (n=27/252)", "Larger (n=109/252)"), 
  small = c("0.35Hg$*$Ni", "0.13Hg$*($Ni$-1)^2$", 
            "0.5Hg$*\\text{race}_{\\text{black}}$", 
            "0.5Hg$*\\text{race}_{\\text{hisp.non}}$"), 
  large = c("0.7Hg$*$Ni", "0.26Hg$*($Ni$-1)^2$",  
            "Hg$*\\text{race}_{\\text{black}}$", 
            "Hg$*\\text{race}_{\\text{hisp.non}}$")
)
labels <- c(
  "Chemical by chemical" = 2, 
  "Chemical by race" = 2
)

equations |> 
  kbl(booktabs = TRUE, escape = FALSE, 
      col.names = c("", "Lower", "Higher"), 
      align = "lcc",
      caption = "Specification of interaction terms in simulations.") |> 
  column_spec(1, width = "10em") |> 
  add_header_above(header = c(" " = 1, "Effect size" = 2)) |>
  pack_rows(index = labels) 
```

# race by ethnicity sensitivity

```{r}
# table
bkmr_re_sens <- read_csv("/Users/elizabethzhang/thesis/thesis/index/data/bkmr_re_sens.csv") |> 
  rename(sensitivity = sens)
oracle_re_int <- read_csv("/Users/elizabethzhang/thesis/thesis/index/data/oracle_re_sens.csv") |> 
  filter(variable == "Int") |> 
  group_by(size, case) |> 
  summarize(sensitivity = sum(p<0.05)/n()) |> 
  mutate(size = ifelse(size == "Small", "Small uncollapsed", size))

re_ints <- bind_rows(
  mutate(bkmr_re_sens, mod = "BKMR"), 
  mutate(oracle_re_int, mod = "Oracle MLR")
) |> 
  arrange(desc(size), desc(mod)) |> 
  pivot_wider(names_from = c(mod, size), values_from = sensitivity) |> 
  mutate(effect_size = ifelse(case %in% c(1, 3), "Lower", "Higher"), 
         case = ifelse(case %in% c(1, 2), 
                       paste0("Original n=27", footnote_marker_symbol(1)), 
                       paste0("Original n=109", footnote_marker_symbol(2)))
         ) |> 
  relocate(case, effect_size) |> 
  select(-5)

re_ints |> 
  kbl(booktabs = TRUE, escape = FALSE, 
      align = "llcccc", 
      caption = "Sensitivity to interactions between the categorical race variable and Hg.", 
      col.names = c("Interaction in", "Effect size", 
                    "Oracle", "BKMR", 
                    "Oracle", "BKMR")
      ) |> 
  add_header_above(header = c(" " = 2, "Small (n=252)" = 2, "Large (n=1000)" = 2), 
                   bold = TRUE) |>
  collapse_rows(columns = 1, valign = "middle", latex_hline = "linespace") |> 
  # column_spec(7, width = "6em") |>
  add_footnote(c("Non-Hispanic black", "Hispanic born outside US"), notation = "symbol", threeparttable = TRUE)
```



# emissions factors for ng by cogen in 2023

```{r}
# read in data 
df <- read_csv("cleaned_02-24.csv")

# get natural gas monthly amounts
ng <- df |> 
  select(month, year, cogen = 4, boilers = 5) |> 
  filter(year != 2024) |> 
  rowwise() |> 
  mutate(total_cf = (cogen + boilers)*100, 
         date = as.Date(paste(year, str_pad(month, 2, pad = "0"), "01", sep = "-")))


comps <- readxl::read_xlsx("naturalgas_emissions.xlsx", sheet = 1)
metals <- readxl::read_xlsx("naturalgas_emissions.xlsx", sheet = 2)

all_poll <- bind_rows(
  mutate(comps, type = "Compounds"), 
  mutate(metals[,2:4], type = "Metals")
)

ng2023 <- ng |> 
  group_by(year) |> 
  summarize(total_cf = sum(total_cf)) |> 
  filter(year == 2023) |> 
  select(-year) |> 
  as.numeric()

all_poll2023 <- all_poll |> 
  janitor::clean_names() |> 
  mutate(emission_factor_lb_106_scf = 
           as.numeric(str_replace_all(emission_factor_lb_106_scf, 
                                      "[,<]", "")), 
         emissions_lb = emission_factor_lb_106_scf * ng2023 / 106)

metals2023 <- all_poll2023 |> 
  filter(type == "Metals")

metals2023 |> 
  select(1, 2, 5, 3) |> 
  mutate(emissions_lb = format(round(emissions_lb, 2), big.mark = ",")) |> 
  kbl(booktabs = TRUE, align = "lccc", escape = FALSE,
      col.names = c("Metal", 
                    linebreak("Factor\n(lb/106scf)", align = "c"), 
                    linebreak("Emissions\n(lb)", align = "c"), "Grade"))

```


